---
title: "GitTyee"
author: "Eric V"
date: "`r Sys.Date()`"
output: html_document
---

A high-level look at trends in catches of Tyee salmon in the legendary Tyee Pool at Campbell River. All of this data exploration has been for [fun](https://i.giphy.com/media/Cz6TlrRVVyv9S/giphy.webp), to learn some new tools and hopefully give myself a leg up in coming years (I need all the help I can get - as do most it seems...). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## Prepare Working Environment ----
## Clear workspace ##
rm(list = ls())  

## This code is to fix a bug in R 2023.06.0+421 <https://github.com/rstudio/rstudio/issues/13188>
options(rstudio.help.showDataPreview = FALSE)
getOption("rstudio.help.showDataPreview")

## Load Packages
    library(tidyverse)
    library(tidyr)
    library(dplyr)
    library(zoo)
    library(lubridate)
    library(reshape2)
    library(ggplot2)
    library(readxl)
    library(downloadthis)
    library(knitr)

## Define Working Directory ---
#setwd("C:/Users/evogt/R Analysis/EAV/Tyee/")

## Define Input Files
    #1. Catch Data
    catch.dat <- "Data/TyeeCatchData.xlsx"
    
    #2. Escapement Data
    esc.dat <- "Data/Area13Escapement.xlsx"
    
    #3. Flow Data
    q.dat <- "Data/CR_Discharge.xlsx"
    
    #4. Age Class Data
    age.dat <- "Data/CR_AgeClass.xlsx"

## Define Output Locations
    plot.output <- "Plots/"
    model.output <- "Output/"
##__________________________####    
## Load Data ---- 

##* 1. Prep Blank Date Data ---- 
    date_seq <- data.frame(Date = seq(ymd('2002-07-01'), ymd('2022-09-30'), by = 'days')) %>%
                mutate(Month = strftime(Date, format = "%m"),
                       Month = as.numeric(Month)) %>%
                filter(between(Month,7,9))

## * 2. Tyee Catch Data ----
    raw_catch  <- read_excel(catch.dat) # Date, time and size of capture

        ## Format Catch Data Dates ##
        raw_catch <- raw_catch %>% mutate(Date = as.Date(raw_catch$RawDate),
                                          catch_binary = 1) %>%
                                   group_by(Date) %>%
                                   summarise(catch_binary = sum(catch_binary))

        ## Merge with date_sequence
        catch_daily <- left_join(date_seq, raw_catch, by = "Date") %>%
                       mutate(catch_binary = coalesce(catch_binary, 0),  
                              Year = strftime(Date, format = "%Y"),
                              Year = as.numeric(Year),
                              Month = strftime(Date, format = "%m"),
                              Month = as.numeric(Month),
                              Weeknum      = strftime(Date, format = "%V"),
                              Weeknum = as.numeric(Weeknum),
                              catch_binary = coalesce(catch_binary, 0))  
      
        
        ## Calculate Yearly Catches
        catch_yearly <- catch_daily %>% 
                        group_by(Year) %>%
                        summarize(catch_yr.total = sum(catch_binary),
                                  catch_yr.mean  = mean(catch_binary)) %>%
                        ungroup %>%
                        mutate(catch_yrs.mean.total = mean(catch_yr.total),
                               catch_yrs.mean.catch = mean(catch_yr.mean),
                               tyee_cycle4 = as.numeric(rep(1:4, length.out = length(Year))),
                               tyee_cycle5 = as.numeric(rep(1:5, length.out = length(Year)))) 
                                          
        ## Calculate Weekly Catches
        catch_yrwk <- catch_daily %>% 
                      group_by(Year, Weeknum) %>%
                      summarize(catch_yrwk.total = sum(catch_binary)) 

        ## Calculate Weekly Catches
        catch_wk <- catch_daily %>% 
                    group_by(Weeknum) %>%
                    filter(between(Weeknum, 27, 39)) %>%
                    summarize(catch_wk.total = sum(catch_binary)) %>%
                    mutate(catch_wk.mean     = catch_wk.total/21) %>%         # 21 years data from 2002 to 2022
                    ungroup() %>%
                    mutate(catch_wks.mean = mean(catch_wk.mean))
        
         
        ## Merge Catch Data
        catch_dat <- left_join(catch_daily, catch_yearly, by = "Year")
        catch_dat <- left_join(catch_dat, catch_yrwk, by = c("Year", "Weeknum"))
        catch_data <- left_join(catch_dat, catch_wk, by = "Weeknum")

        colnames(catch_data)                               
 
## * 3. Escapement Data ----
        ## Load Data      
        esc_data <- read_excel(esc.dat)
    
        ## Prepare dataset for CR Escapement 
        cr_esc <- esc_data %>% filter(GAZETTED_NAME == "CAMPBELL RIVER",                 # only show data for CR
                                      ANALYSIS_YR >=2002) %>%                             # data for same time period as tyee catch data    
                               mutate(Year          =  as.numeric(ANALYSIS_YR),
                                     Waterbody      =  as.factor(GAZETTED_NAME),
                                     Esc_Count      =  as.numeric(MAX_ESTIMATE),
                                     Esc_Scaled     =  as.numeric(MAX_ESTIMATE/25),
                                     Esc_Mean       =  as.numeric(mean(MAX_ESTIMATE)),
                                     Esc_MeanScaled =  as.numeric(mean(MAX_ESTIMATE/25)),
                                     Species        =  as.factor(SPECIES)) %>%
                               select(Year, Waterbody, Species, Esc_Count, Esc_Scaled, Esc_Mean, Esc_MeanScaled)
        
        
        
## * 4. Flow Data ----
    flow_data <- read_excel(q.dat)

    ## Calculate Weekly Flows    
    flow_data <- flow_data %>% 
                 filter(between(Month, 7, 9), Year >= 2002) %>% 
                        mutate(q.daily = as.numeric(Value),
                               Weeknum = strftime(Date, format = "%V"), #create "Weeknum" column
                               Weeknum = as.numeric(Weeknum),
                               Date    = as.Date(Date),
                               q.5yr.roll.mean = rollmean(q.daily, k = 455, align = "right", fill=NA), # 5yr running flow
                               q.5y.roll.se = rollapply(q.daily, width = 455, FUN = sd, 
                                                        align = "right", fill = NA)/sqrt(455))  # 5yr running SE
              str(flow_data)
    
    q.yearly <- flow_data %>% group_by(Year) %>% 
                                 summarize(q.yr.mean   = mean(Value),
                                           q.yr.se     = sd(Value)/sqrt(n()))
    
    q.monthly <- flow_data %>% group_by(Year, Month) %>% 
                                  summarize(q.yrm    = mean(Value),
                                            q.yrm.se = sd(Value)/sqrt(n()))
    
    q.m       <- flow_data %>% group_by(Month) %>% 
                                  summarize(q.m    = mean(Value),
                                            q.m.se = sd(Value)/sqrt(n()))
      
    q.weekly <- flow_data %>% group_by(Year, Weeknum) %>% 
                                 summarize(q.yrwk    = mean(Value),
                                           q.yrwk.max = max(Value),
                                           q.yrwk.se = sd(Value)/sqrt(n()))
    
    q.wk <-flow_data %>% group_by(Weeknum) %>%
                                summarize(q.wk    = mean(Value),
                                          q.wk.se = sd(Value)/sqrt(n()))
                                
    ## Merge Flow Datasets
    flow.dat <- left_join(flow_data, q.yearly, by = "Year")
    flow.dat <- left_join(flow.dat, q.monthly, by = c("Year","Month"))
    flow.dat <- left_join(flow.dat, q.m,       by = ("Month"))
    flow.dat <- left_join(flow.dat, q.wk,      by = ("Weeknum"))
    flow.data <- left_join(flow.dat, q.weekly, by = c("Year", "Weeknum"))
    
    
## 5. Add age data
    age_data <- read_excel(age.dat)

```

## The Data

This analysis uses the following publicly available data:

* Annual catch records from the [Tyee Club](https://www.tyeeclub.org/catch-records/>). 
* Discharge data collected on the Campbell River by the [Water Survey of Canada](https://wateroffice.ec.gc.ca/index_e.html).
* Annual Chinook Salmon escapement data available in the [DFO NuSEDS database](https://open.canada.ca/data/en/dataset/c48669a3-045b-400d-b730-48aafe8c5ee6).
  

This is certainly an interesting dataset, but it has its limitations. For example, there is no information on effort (# boats per day), age structure of captured Tyees or numbers of undersize Chinook captured in the pool.

For now, I have only looked at data back to 2002, as that is the limit of easily accessible Tyee catch data. 

## TyeePlot  
```{r TyeeCatches, warnings = FALSE, fig.cap="Figure 1. Number of Tyee Salmon captured per year."}
## Prepare Plot
    ggplot(catch_data) +
            geom_bar(aes(x= Year, y = catch_binary, fill = tyee_cycle4), stat = "identity") +
            geom_hline(aes(yintercept=catch_yrs.mean.total, color = catch_yrs.mean.total), 
                      linetype = "dashed", linewidth = 0.5) +
            ylab("# Tyee Salmon") +
            guides(fill = "none", #guide_legend(title = "Tyee Cycle"),
                   color = guide_legend(title = "Mean catches per year", label = FALSE)) + 
            scale_x_continuous(breaks = seq(2002,2023,2)) +
            scale_y_continuous(breaks = seq(0, 100, 5)) +
            theme_bw() +  
            theme(legend.position = c(0.78, 0.88),
                  legend.box.background = element_rect(colour = "black", linewidth = 1.25, fill = NA)) 

```



A quick look at the figure shows:
  1. There appears to be a 4-year cycle of high catches (highlighted with shading). 
    * Publically accessible information on the age structure of Chinook Salmon in the Campbell River is relatively limited. However,               [Sturham et al. 1999](https://publications.gc.ca/collections/collection_2014/mpo-dfo/Fs97-4-2477-eng.pdf) report that fish returning         to the Campbell River are primarily Age 4 (length from 550-949 mm, mean = 780 mm) and Age 5 (length = 700-949 mm, mean = 846 mm).            Tyees captured in the Tyee pool are typically over XXX mm. More recent information on the age composition of Campbell River Chinook          likely exists, but I haven't found it. 

```{r table prep, include = FALSE}
age_table <- age_data %>% select(Waterbody, Sex, Age, n, Perc.Total.by.Sex, Perc.Total.by.Stream, bin_range, meanFL, SE)

```
      
        
```{r age table}

knitr::kable(age_table, col.names = c("Waterbody","Sex", "Age", "n", "% of Total <br>(by Sex)", "% of Total <br>(by Stream)", "Size Range <br>(mm)", "Mean Lenght <br>(mm)", "Std. Error"),
             digits = 2,  align=rep('c', 5),
             caption = "Table 1. Size at age of Chinook Salmon captured in Campbell River watershed")
```

  2. There has been a consistent decline in the number of Tyee salmon captured per year and there was a major crash in 2014.IF we look at         Figure 2, we can see:
      * There has been a general declining trend in escapement,
      * Periods of increased escapement correspond with periods of increased Tyee catches, but not always (e.g. 2005, 2017 and 2020). Given          the lack of effort data we cannot tease apart whether the lack of catches in some years is due to reduction in pressure. 
      * It is also possible that years with high escapement and low Tyee numbers were due to an increased proportion of smaller fish                 returning to the Campbell. Without annual information on age structure I cannot tease this apart.
    
```{r include=FALSE}
## Merge escapement data with tyee catch data
cr_data <- left_join(catch_data, cr_esc, by = "Year")
    
cr_yearly <- cr_data %>% mutate(Year = as.numeric(Year)) %>%
                         group_by(Year) %>%
                         select(Year, catch_yr.total, catch_yr.mean, Esc_Scaled, Species, tyee_cycle4)
```

## Tyee Escapement Plot        
```{r fig_prep}

## Prepare plot
cr_plot<- ggplot(cr_data) +
                geom_bar(aes(x = Year, y = catch_binary, fill = tyee_cycle4), 
                              stat = "identity") +
                geom_line(aes(x = Year, y = Esc_Scaled, linetype = Species), 
                              color = "lightblue", linewidth = 1.2) +
                geom_hline(aes(yintercept = catch_yrs.mean.total, color = catch_yrs.mean.total), 
                              linetype = "dashed", linewidth = 0.5) +
                labs(x = "Year", y = "Number Tyee Captured") + 
                guides(fill =  "none", #guide_legend(title = "Tyee Cycle", label = FALSE),
                       linetype = guide_legend(title = "Escapement", label = FALSE),
                       color = guide_legend(title = "Mean Tyee Catch", label = FALSE)) +
                scale_y_continuous(breaks = seq(0, 100, 5),
                                    sec.axis = sec_axis(~.*25, name = "Escapement")) +  #scale secondary access by x25
                scale_x_continuous(breaks = seq(2002,2022,2)) +            
                theme_bw() +
                theme(legend.direction = "horizontal",
                      legend.position = c(0.78, 0.88),
                      legend.box.background = element_rect(colour = "black", linewidth = 1.25, fill = "white"))

```

```{r tyee vs escapement, fig.cap = "Figure 2. Trends in Tyee Salmon captures and Campbell River Chinook Salmon escapement", echo=FALSE}
suppressWarnings(print(cr_plot))   
```


Let's look at flow in the Campbell River to see how environmental conditions may influence catches.
```{r data prep catch_flow, include = FALSE}
    flow.catch <- left_join(flow.data, catch_data, by = c("Year","Month","Weeknum","Date"))
    
    
    ## Reduce to past 5 years only.                          
    flow.catch.recent <- flow.catch %>%  
                         mutate(catch_scaled    = catch_binary * 7, 
                                q.daily.scaled  = q.daily/20,
                                q.5yr.rm.scaled = q.5yr.roll.mean/20,
                                q.5y.rse.scaled = q.5y.roll.se/20,
                                date.std = case_when(year(Date) >= 0 ~ 'year<-'(Date, 2021), ## Set all dates to same year
                                                 TRUE ~ Date)) %>%
                                filter(between(date.std, as.Date("2021-07-15"), as.Date("2021-09-15")),
                                               Year >2010,
                                               PARAM == 1)
```

```{r plot flow vs catches, fig.dim = c(8,16), warnings = FALSE, message = FALSE}
## Prepare plot    
plot <-    ggplot(flow.catch.recent) +
                 geom_bar(aes(x = date.std, y = catch_scaled, fill = PARAM), stat= "identity", color = "Navy") +
                 geom_line(aes(x = date.std, y = q.daily, linetype = "dashed", color = "Daily")) +
                 geom_line(aes(x = date.std, y = q.5yr.roll.mean, linetype = "twodashed", color = "5-yr Mean")) +
                 labs(x = "Date", y = "Discharge") +
                 scale_x_date(date_breaks = "5 days", date_labels = "%m-%d", minor_breaks = "1 day") +
                 scale_y_continuous(breaks = seq(0, 75, 10),
                                    sec.axis = sec_axis(~./7, name = "Number of Fish")) +  #scale secondary access by x25
                 guides(fill = guide_legend(title = "# Tyee", label = FALSE),
                        linetype = "none",
                        color = guide_legend(title = "Discharge (cms)", label = TRUE)) +
                 facet_grid(Year~.) +
                 theme_bw() +
                 theme(legend.position = "bottom")  
suppressWarnings(print(plot))
```

